---
title: "Data Science 1: Homework 1"
author: Caroline Andy
date: 9/18/2020
output: github_document
---

Here are my solutions to homework 1. 

```{r tidyverse, message = FALSE, results = 'hide', warning = FALSE}
library(tidyverse)
```

**Problem 1**

Generate a data frame containing the required elements.

```{r create df}
prob1_df = 
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a","b","c","d","e","f","g","h","i","j"),
    factor_vec = factor(c("level_1","level_1","level_1","level_2","level_2","level_2","level_3","level_3","level_3","level_3")))
```

_Calculating mean_

Next I will attempt to take the mean of each variable.

```{r mean, warning = FALSE}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```

I am able to take the mean of my numerical and logical variables, samp and samp_gt_0, but not my character and factor variables, char_vec and factor_vec.

_Converting variable types_

In some cases, I can explicitly convert variables from one type to another. Here, I will attempt to convert each of my variables to numeric. 

```{r try numeric, warning = FALSE, results = 'hide'}
as.numeric(pull(prob1_df, samp))
as.numeric(pull(prob1_df, samp_gt_0))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, factor_vec))
```

The as.numeric function generates numeric variable outputs for the numeric and logical vectors, samp and samp_gt_0. The samp field, numeric to begin with, remained unchanged. The mean of samp, generated above, was calculated using its numeric entries. 

The samp_gt_0 field was a logical vector with 0 and 1 values denoting whether the samp entries are above or below 0. This binary field was converted into numeric using the as.numeric function. In the above example, the mean function generated the mean for samp_gt_0 by first converting its logical entries into numeric values. 

The character vector was unable to be converted into a numeric field; we cannot calculate a mean for this variable. 

The factor vector was able to be coerced into a numeric variable; however, the mean cannot be calculated until the variable is first manually converted into a numeric variable. 

_Multiplying coerced variables_

Now I will multiply the random sample variable by the logical vector. 
```{r as.numeric multiplication, warning = FALSE}
as.numeric(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp) 
```

Now I will convert the logical vector to a factor, and multiply the random sample by the result.
```{r as.factor multiplication, warning = FALSE}
as.factor(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp) 
```

Now I will convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result
```{r more multiplication, warning = FALSE}
as.numeric(as.factor(pull(prob1_df, samp_gt_0))) * pull(prob1_df, samp) 
```

**Problem 2**

_Loading and describing dataset_

For the following problem, I will load the penguins dataset. 

```{r load data}
data("penguins", package = "palmerpenguins")
```

The penguin dataset contains `r nrow(penguins)` rows, where each row corresponds with an observed penguin. This dataset contains `r ncol(penguins)` columns, which include the following variables: penguin species, island, bill length, bill depth, flipper length, body mass, sex and year. 

Data are collected from years `r min(as.numeric(pull(penguins, year)))` to `r max (as.numeric(pull(penguins,year)))`.
There are `r nlevels(as.factor(pull(penguins, species)))` different species (`r levels(as.factor(pull(penguins, species)))`), and `r nlevels(as.factor(pull(penguins, island)))` different islands (`r levels(as.factor(pull(penguins, island)))`) represented in the penguins dataset.

The mean bill length is `r round(mean(na.omit(pull(penguins, bill_length_mm))), digits = 2)` millimeters. Mean body mass is `r round(mean(na.omit(pull(penguins, body_mass_g))), digits = 2)` grams. The mean flipper length is `r round(mean(na.omit(pull(penguins, flipper_length_mm))), digits = 2)` millimeters.

_Generating a scatterplot_ 

Next I will generate a scatterplot graphing bill length and flipper length for each observed penguin, where dot color is determined by penguin species. 

```{r scatterplot, warning = FALSE}
# generating scatterplot
library(ggplot2)
penguinsggplot <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
#viewing scatterplot
penguinsggplot
# exporting scatterplot
ggsave("penguinsggplot.pdf")
```